### 2014-03-18 presentation.Rnw
Make a few slides!

\documentclass[12pt]{beamer}

\title{Survival Analysis}
\subtitle{Senior Statistics Seminar}
\author{Brian To}

\begin{document}
\SweaveOpts{concordance=TRUE}
  \maketitle

  %%% Introduction
  \begin{frame}{Sounds cool\ldots but what is it?}
    \begin{itemize}
      \item What proportion will ``survive'' past a certain time?
      % Event History Analysis (sociology):
      %   Which kind of strikes last more than 3 days?
      \item Of those that ``survive,'' what is their failure or death rate?
      % Reliability Analysis (engineering):
      %   How long do different types of F1 tyres last?
      \item What factors affect how individuals ``survive''?
      % Duration Modeling (economics):
      %   How do different factors affect unemployment duration?
    \end{itemize}
  \end{frame}

  %%% History
  \begin{frame}{How did this magic happen?}
    \begin{itemize}
      \item Analyze actuarial data
      % for mortality tables, very informal
      \item Became a discipline in WWII
      % interest reliability of equipment
      % spurred on by spike in demands for reliability, quality
      \item Medical field drove much of survival analysis
      % b/c studies do not need all subjects at the start of the study (async)
      %   ...handles subjects that quit, cannot follow-up, or die of different event
    \end{itemize}
  \end{frame}
\end{document}

### 2014-03-21 worksheet.R
Try to start worksheet

@@ -1,0 +1,10 @@
+durations <- 1 + rpois(n = 10, lambda = 3)
+offsets <- sample(0:5, size = 10, replace = TRUE)
+ends <- offsets + durations
+event <- sapply(ends, function(end) { return(end < 8) })
+timeline <- data.frame(start = offsets, end = ends, event = event)
+
+# Study ends at t = 7
+survival <- survfit(Surv(start, end, event) ~ 1, data = timeline)
+
+timeline.disp <- with(timeline, )

### 2014-03-28 presentation.Rnw
Add notes and things
Retry slides!

@@
   \maketitle
+
+  \begin{frame}{Introduction}
+    \begin{center}
+      What proportion of the population will\\
+      \emph{survive past a given time}?
+    \end{center}
+  \end{frame}
+
+  % http://rbakker.myweb.uga.edu/pols8501/OxfordOneNotes.pdf
+  \begin{frame}{Applications}
+    \begin{itemize}
+      \item Medical -- How does flu survival differ between countries?
+      \item Sociology -- What kind of strikes last more than 3 days?
+      \item Engineering -- How long do different F1 tyres last?
+      \item Economics -- What affect unemployment duration?
+      \item Politics -- How long are civil wars? peaceful times?
+      \item Actuarial Science -- How long do people live in cities?
+    \end{itemize}
+  \end{frame}
+
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%
+  % Old Stuff 
+  %%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
     \end{itemize}
+
+    % also: political careers, earthquakes, medical, 
   \end{frame}
@@
   %%% History
+  %%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
+  %%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
   \begin{frame}{How did this magic happen?}
@@
       %   ...handles subjects that quit, cannot follow-up, or die of different event
+      %   (handles incomplete data)
+      \item Kicked off with Kaplan-Meier estimator
+      % http://www.jehps.net/juin2009/Aalenetal.pdf
+      % raises new questions: how to compare survival curves?
+      % prop hazards resolved covariates problem
+      % application was ahead of theory, didn't know why things worked
     \end{itemize}

### 2014-03-29 presentation.Rnw
Reworked introduction
Formatting, worksheet plots timeline

@@
 
+% http://www.nytimes.com/2011/08/13/health/13meier.html
 \begin{frame}{Introduction}
@@
     \emph{survive past a given time}?
+  \end{center}
+
+  \vspace{0.2in}
+  \pause
+
+  \begin{center}
+    We are measuring the \emph{time to event}
+  \end{center}
+
+  \vspace{0.2in}
+  \pause
+
+  \begin{center}
+    How do flu durations differ between countries?
   \end{center}
@@
   \begin{itemize}
-      \item Medical -- How does flu survival differ between countries?
+    % Event History Analysis
     \item Sociology -- What kind of strikes last more than 3 days?
+    % Reliability Analysis
     \item Engineering -- How long do different F1 tyres last?
+    % Duration Analysis
     \item Economics -- What affect unemployment duration?
     \item Politics -- How long are civil wars? peaceful times?
-      \item Actuarial Science -- How long do people live in cities?
   \end{itemize}
+\end{frame}
+
+\begin{frame}{Data Collection Issues}
+<<example-timeline, echo=FALSE, fig=TRUE>>=
+
+@
 \end{frame}

### 2014-03-29 worksheet.R

@@ -7,4 +7,6 @@ timeline <- data.frame(start = offsets, end = ends, event = event)
 # Study ends at t = 7
 survival <- survfit(Surv(start, end, event) ~ 1, data = timeline)
 
-timeline.disp <- with(timeline, )
+timeline.plot <-
+  ggplot(timeline) +
+  geom_segment(aes(x = start, xend = end, y = rownames(timeline), yend = rownames(timeline)))

### 2014-03-29 presentation.Rnw
Got sample timeline up!

@@
   \maketitle
+
+<<echo=FALSE>>=
+library(ggplot2)
+library(survival)
+@
 
@@
 \begin{frame}{Data Collection Issues}
-<<example-timeline, echo=FALSE, fig=TRUE>>=
+<<example-timeline, echo=FALSE, fig=TRUE, width=5, height=4>>=
+timeline <- read.csv('example-timeline.csv', header=TRUE, row.names=1)
 
+ggplot(timeline) +
+  xlab('time') +
+  ylab('observation') +
+  labs(
+    colour = "Observation",
+    shape = "Event Occured?")+
+  geom_segment(aes(
+    colour = rownames(timeline),
+    x = start,
+    xend = end,
+    y = rownames(timeline),
+    yend = rownames(timeline))) +
+  geom_point(aes(
+    colour = rownames(timeline),
+    x = start,
+    y = rownames(timeline)), size = 4, title="asdf") +
+  geom_point(aes(
+    colour = rownames(timeline),
+    shape = event,
+    x = end,
+    y = rownames(timeline)), size = 4) +
+  geom_vline(
+    xintercept = 6,
+    colour = 'red',
+    linetype = 'longdash')
 @

### 2014-03-29 worksheet.R

@@ -1,12 +1,32 @@
-durations <- 1 + rpois(n = 10, lambda = 3)
-offsets <- sample(0:5, size = 10, replace = TRUE)
-ends <- offsets + durations
-event <- sapply(ends, function(end) { return(end < 8) })
-timeline <- data.frame(start = offsets, end = ends, event = event)
+timeline <- read.csv('example-timeline.csv', header=TRUE, row.names=1)
 
 # Study ends at t = 7
 survival <- survfit(Surv(start, end, event) ~ 1, data = timeline)
 
-timeline.plot <-
-  ggplot(timeline) +
-  geom_segment(aes(x = start, xend = end, y = rownames(timeline), yend = rownames(timeline)))
+timeline.plot <- ggplot(timeline) +
+  xlab('time') +
+  ylab('observation') +
+  labs(
+    colour = "Observation",
+    shape = "Event Occured?")+
+  geom_segment(aes(
+    colour = rownames(timeline),
+    x = start,
+    xend = end,
+    y = rownames(timeline),
+    yend = rownames(timeline))) +
+  geom_point(aes(
+    colour = rownames(timeline),
+    x = start,
+    y = rownames(timeline)), size = 4, title="asdf") +
+  geom_point(aes(
+    colour = rownames(timeline),
+    shape = event,
+    x = end,
+    y = rownames(timeline)), size = 4) +
+  geom_vline(
+    xintercept = 6,
+    colour = 'red',
+    linetype = 'longdash')
+
+timeline.plot

### 2014-03-30 presentation.Rnw
Fix screwups

--- 2014-03-29 presentation.Rnw
@@
 \begin{frame}{Data Collection Issues}
-<<example-timeline, echo=FALSE, fig=TRUE>>=
+<<example-timeline, echo=FALSE, fig=TRUE, width=5, height=4>>=
+timeline <- read.csv('example-timeline.csv', header=TRUE, row.names=1)
 
+ggplot(timeline) +
+  xlab('time') +
+  ylab('observation') +
+  labs(
+    colour = "Observation",
+    shape = "Event Occured?")+
+  geom_segment(aes(
+    colour = rownames(timeline),
+    x = start,
+    xend = end,
+    y = rownames(timeline),
+    yend = rownames(timeline))) +
+  geom_point(aes(
+    colour = rownames(timeline),
+    shape = event,
+    x = end,
+    y = rownames(timeline)), size = 4) +
+  geom_vline(
+    xintercept = 6,
+    colour = 'red',
+    linetype = 'longdash')
 @

### 2014-03-31 presentation.Rnw
Added descriptors slide
Added example graphs and confused myself
hazards is a rate, but of what to what?

--- 2014-03-29 presentation.Rnw
@@
   \maketitle
+
+<<echo=FALSE>>=
+library(ggplot2)
+library(survival)
+library(KMsurv) # survival data sets
+library(GGally) # survival curve support for ggplot2
+
+data(bfeed)
+
+# FDA recommends 6 months of breastfeeding ~ 27 weeks
+bfeed.short <- subset(bfeed, duration < 32)
+@
 
@@
 \begin{frame}{Data Collection Issues}
-<<example-timeline, echo=FALSE, fig=TRUE>>=
+\begin{center}
+<<example-timeline, echo=FALSE, fig=TRUE, height=4, width=6>>=
+timeline <- read.csv('example-timeline.csv', header=TRUE, row.names=1)
 
+ggplot(timeline) +
+  ggtitle('Example Event Timeline') +
+  xlab('time') +
+  ylab('observation') +
+  labs(
+    colour = "Observation",
+    shape = "Event Occured?")+
+  geom_segment(aes(
+    colour = rownames(timeline),
+    x = start,
+    xend = end,
+    y = rownames(timeline),
+    yend = rownames(timeline))) +
+  geom_point(aes(
+    colour = rownames(timeline),
+    shape = event,
+    x = end,
+    y = rownames(timeline)), size = 4) +
+  geom_vline(
+    xintercept = 6,
+    colour = 'red',
+    linetype = 'longdash')
 @
+\end{center}
+\end{frame}
+
+\begin{frame}{Descriptors}
+Assume $f(t)$ is the instantaneous failure rate
+
+\vspace{0.5in}
+
+% Hazard stuff is wrong
+\begin{tabular}{lllcc}
+Survival
+  & $P(\mbox{no event in } t_i < t)$
+  & $S(t)$ & = & $\int^{t}_{0} f(x)$ \\
+Hazard
+  & $P(\mbox{no event until } t)$
+  & $\lambda(t)$ & = & $^{f(t)} / _{S(t)}$ \\
+Cum. \\ Hazard
+  & E()
+  & $\Lambda(t)$ & = & $\int^{t}_{0} \lambda(t) dt$ \\
+\end{tabular}
+\end{frame}
+
+\begin{frame}{Descriptors}
+\begin{tabular}{lcc}
+  & independent & conditional \\
+  instantaneous & $f(t)$ & $\lambda(t)$ \\
+  cumulative & $S(t)$ & $\Lambda(t)$
+\end{tabular}
+\end{frame}
+
+\begin{frame}{Example}
+\begin{center}
+<<bfeed-plain, echo=FALSE, fig=TRUE, height=4, width=6>>=
+bfeed.fit <- survfit(Surv(duration, delta) ~ 1, data = bfeed.short)
+ggsurv(bfeed.fit) +
+  ggtitle('Breast Feeding Duration') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+@
+\end{center}
+\end{frame}
+
+\begin{frame}{Models}
+(Cox) Proportional Hazards Model
+
+$$\lambda(t | \mathbf{x}) = \lambda_{0}(t) \exp \left( \mathbf{x' \beta} \right)$$
+
+Accelerated Failure Time Model
+
+$$\log T = \mathbf{x' \beta} + \mathbf{\epsilon}$$
+
+Exponential/Weibull Model
+
+$$\lambda(t, \mathbf{x}) = \lambda_{0} \exp \left( \mathbf{x' \beta} \right)$$
 \end{frame}

### 2014-03-31 presentation.Rnw
Figured out hazards function
It's an odds ratio, not a probability. Internets are so fucking wrong.

@@
 \begin{frame}{Descriptors}
-Assume $f(t)$ is the instantaneous failure rate
+Assume $f(t)$ is the instantaneous probability of the event occurring.
 
@@
 
-% Hazard stuff is wrong
 \begin{tabular}{lllcc}
 Survival
-  & $P(\mbox{no event in } t_i < t)$
-  & $S(t)$ & = & $\int^{t}_{0} f(x)$ \\
+  & $P(\mbox{no event before } t)$
+  & $S(t)$ & = & $1 - \int^{t}_{0} f(x)$ \\
 Hazard
-  & $P(\mbox{no event until } t)$
+  & $OR(\mbox{no event until } t)$
   & $\lambda(t)$ & = & $^{f(t)} / _{S(t)}$ \\
-Cum. \\ Hazard
-  & E()
+Cumulative \\ Hazard
+  & $OR(\mbox{event before } t)$
   & $\Lambda(t)$ & = & $\int^{t}_{0} \lambda(t) dt$ \\
@@
 \begin{tabular}{lcc}
-  & independent & conditional \\
+  & probability & odds ratio \\
   instantaneous & $f(t)$ & $\lambda(t)$ \\

### 2014-03-31 presentation.Rnw
Add hazard function graph and conf bands.

@@
 \begin{center}
-<<example-timeline, echo=FALSE, fig=TRUE, height=4, width=6>>=
-timeline <- read.csv('example-timeline.csv', header=TRUE, row.names=1)
+<<example-timeline, echo=FALSE, fig=TRUE, height=4>>=
+timeline <- read.csv('example-timeline.csv', header = TRUE, row.names = 1)
 
@@
 \begin{center}
-<<bfeed-plain, echo=FALSE, fig=TRUE, height=4, width=6>>=
+<<bfeed-survival, echo=FALSE, fig=TRUE, height=4>>=
 bfeed.fit <- survfit(Surv(duration, delta) ~ 1, data = bfeed.short)
@@
   geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+@
+\end{center}
+\end{frame}
+
+\begin{frame}{Example}
+\begin{center}
+<<bfeed-hazard, echo=FALSE, fig=TRUE, height=4>>=
+# bfeed.haz <- basehaz(coxph(Surv(duration, delta) ~ 1, data = bfeed.short))
+bfeed.haz <- with(bfeed.fit, -log(data.frame(
+  hazard = surv, # haz = -log(surv)
+  upper = upper,
+  lower = lower)))
+bfeed.haz$week <- as.numeric(rownames(bfeed.haz))
+
+ggplot(bfeed.haz, aes(x = week)) +
+  geom_step(aes(y = hazard)) +
+  geom_step(aes(y = lower),
+    colour = 'black',
+    linetype = 'dashed') +
+  geom_step(aes(y = upper),
+    colour = 'black',
+    linetype = 'dashed') +
+  ggtitle('Breast Feeding Duration') +
+  xlab('Duration in Weeks') +
+  ylab('Cumulative Hazard Ratio') +
+  geom_vline(aes(xintercept = 25),
     colour = 'red',

### 2014-04-01 worksheet.R
Add formulas, make class activity

@@ -1,32 +1,35 @@
-timeline <- read.csv('example-timeline.csv', header=TRUE, row.names=1)
+ws <- read.csv('worksheet.csv', header = TRUE)
+ws$id <- sprintf('%02d', 1:nrow(ws))
 
-# Study ends at t = 7
-survival <- survfit(Surv(start, end, event) ~ 1, data = timeline)
-
-timeline.plot <- ggplot(timeline) +
+ws.rawdisp <- ggplot(ws, aes(y = id, yend = id, colour = event)) +
+  ggtitle('Event Timeline') +
   xlab('time') +
   ylab('observation') +
-  labs(
-    colour = "Observation",
-    shape = "Event Occured?")+
   geom_segment(aes(
-    colour = rownames(timeline),
     x = start,
     xend = end,
-    y = rownames(timeline),
-    yend = rownames(timeline))) +
+    order = id)) +
+  geom_vline(aes(
+    xintercept = max(end)), colour = 'red', linetype = 'longdash') +
   geom_point(aes(
-    colour = rownames(timeline),
-    x = start,
-    y = rownames(timeline)), size = 4, title="asdf") +
-  geom_point(aes(
-    colour = rownames(timeline),
-    shape = event,
-    x = end,
-    y = rownames(timeline)), size = 4) +
-  geom_vline(
-    xintercept = 6,
-    colour = 'red',
-    linetype = 'longdash')
+    x = end), size = 4)
 
-timeline.plot
+ws.durationdisp <- ggplot(ws, aes(y = id, yend = id, colour = event)) +
+  ggtitle('Event Duration') +
+  xlab('duration') +
+  ylab('observation') +
+  geom_segment(aes(
+    x = 0,
+    xend = duration,
+    order = id)) +
+  geom_point(aes(
+    x = duration), size = 4)
+
+ggblank <- ggplot(data.frame()) + geom_point()
+
+ans.failure <- ggblank +
+  xlim(0, max(ws$duration)) +
+  ylim(0, 1) +
+  scale_y_continuous(
+    breaks = 0:9 / 9,
+    minor_breaks = 0:18 / 18)

### 2014-04-01 worksheet.R
worksheet.R no longer needed

@@ -1,35 +1,0 @@
-ws <- read.csv('worksheet.csv', header = TRUE)
-ws$id <- sprintf('%02d', 1:nrow(ws))
-
-ws.rawdisp <- ggplot(ws, aes(y = id, yend = id, colour = event)) +
-  ggtitle('Event Timeline') +
-  xlab('time') +
-  ylab('observation') +
-  geom_segment(aes(
-    x = start,
-    xend = end,
-    order = id)) +
-  geom_vline(aes(
-    xintercept = max(end)), colour = 'red', linetype = 'longdash') +
-  geom_point(aes(
-    x = end), size = 4)
-
-ws.durationdisp <- ggplot(ws, aes(y = id, yend = id, colour = event)) +
-  ggtitle('Event Duration') +
-  xlab('duration') +
-  ylab('observation') +
-  geom_segment(aes(
-    x = 0,
-    xend = duration,
-    order = id)) +
-  geom_point(aes(
-    x = duration), size = 4)
-
-ggblank <- ggplot(data.frame()) + geom_point()
-
-ans.failure <- ggblank +
-  xlim(0, max(ws$duration)) +
-  ylim(0, 1) +
-  scale_y_continuous(
-    breaks = 0:9 / 9,
-    minor_breaks = 0:18 / 18)

### 2014-04-01 presentation.Rnw
Add formulas, make class activity
Fix up kaplan-meier slide, add n_i formula
Fixed desc. formatting, removed old stuff

@@
 
+\renewcommand{\arraystretch}{1.5}
 \begin{tabular}{lllcc}
@@
   & $\lambda(t)$ & = & $^{f(t)} / _{S(t)}$ \\
-Cumulative \\ Hazard
+Cum. Haz.
   & $OR(\mbox{event before } t)$
   & $\Lambda(t)$ & = & $\int^{t}_{0} \lambda(t) dt$ \\
+  & & & = & $- \log{S(t)}$ \\
 \end{tabular}
@@
 
+\begin{frame}{Estimators}
+Kaplan-Meier Max Likelihood Estimator for $\hat S(t)$
+
+\vspace{0.2in}
+\begin{tabular}{ll}
+  $t_i$ & time interval \\
+  $d_i$ & \# events in $t_i$ \\
+  $n_i$ & \# subjects \emph{at risk} in $t_i$ \\
+\end{tabular}
+
+\begin{eqnarray*}
+  \hat S(t_0) & = & 1 \\
+  \hat S(t)   & = & \prod_{t_i < t} \frac{n_i - d_i}{n_i} \\
+              & = & \frac{n_i - d_i}{n_i} \cdot S(t_{i - 1}) \\ \\
+    n_{i + 1} & = & n_i - d_i - c_i
+\end{eqnarray*}
+\end{frame}
+
 \begin{frame}{Models}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Old Stuff 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% Introduction
-\begin{frame}{Sounds cool\ldots but what is it?}
-  \begin{itemize}
-    \item What proportion will ``survive'' past a certain time?
-    % Event History Analysis (sociology):
-    %   Which kind of strikes last more than 3 days?
-    \item Of those that ``survive,'' what is their failure or death rate?
-    % Reliability Analysis (engineering):
-    %   How long do different types of F1 tyres last?
-    \item What factors affect how individuals ``survive''?
-    % Duration Modeling (economics):
-    %   How do different factors affect unemployment duration?
-  \end{itemize}
-
-  % also: political careers, earthquakes, medical, 
-\end{frame}
-
-%%% History
 %%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
 %%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
-\begin{frame}{How did this magic happen?}
-  \begin{itemize}
-    \item Analyze actuarial data
-    % for mortality tables, very informal
-    \item Became a discipline in WWII
-    % interest reliability of equipment
-    % spurred on by spike in demands for reliability, quality
-    \item Medical field drove much of survival analysis
-    % b/c studies do not need all subjects at the start of the study (async)
-    %   ...handles subjects that quit, cannot follow-up, or die of different event
-    %   (handles incomplete data)
-    \item Kicked off with Kaplan-Meier estimator
-    % http://www.jehps.net/juin2009/Aalenetal.pdf
-    % raises new questions: how to compare survival curves?
-    % prop hazards resolved covariates problem
-    % application was ahead of theory, didn't know why things worked
-  \end{itemize}
-\end{frame}
+%%% http://www.jehps.net/juin2009/Aalenetal.pdf
 \end{document}

### 2014-04-02 presentation.Rnw
Fix stuff
Add dataset graphs and tweak equations

@@
 library(ggplot2)
+library(gridExtra) # multi-plot graphs
 library(survival)
@@
 library(GGally) # survival curve support for ggplot2
+library(xtable) # make latex tables
 
@@
 
-\begin{frame}{Example}
+\begin{frame}{Dataset}
+\begin{itemize}
+  \item National Labor Survey of Youth
+  \item Examined pregnancies before 1983
+  \item Restricted to first-born child
+  \item Restricted to gestation of between 20 and 45 weeks
+  \item Restricted to subjects with complete responses
+\end{itemize}
+
+\tiny
+<<bfeed-table, echo = FALSE, results = TEX>>=
+bfeed.ids <- sample(rownames(bfeed.short), size = 8)
+bfeed.table <- bfeed.short[bfeed.ids, ]
+
+print(xtable(bfeed.table))
+@
+\end{frame}
+
+\begin{frame}{Dataset}
 \begin{center}
@@
 
-\begin{frame}{Example}
+\begin{frame}{Dataset}
 \begin{center}
@@
 
-\begin{frame}{Estimators}
-Kaplan-Meier Max Likelihood Estimator for $\hat S(t)$
+\begin{frame}{Dataset}
+\begin{center}
+<<bfeed-by-education, echo = FALSE, fig = TRUE, height = 4>>=
+bfeed.short$college <- bfeed.short$yschool > 12
+bfeed.fit <- survfit(Surv(duration, delta) ~ college, data = bfeed.short)
 
-\vspace{0.2in}
+ggsurv(bfeed.fit) +
+  ggtitle('Breast Feeding Duration by Education') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+@
+\end{center}
+\end{frame}
+
+\begin{frame}{Dataset}
+\begin{center}
+<<bfeed-by-poverty, echo = FALSE, fig = TRUE, height = 5>>=
+bfeed.fit <- survfit(Surv(duration, delta) ~ poverty, data = bfeed.short)
+
+bfeed.poverty.short <- ggsurv(bfeed.fit) +
+  ggtitle('Breast Feeding Duration by Poverty, Duration < 30 Weeks') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+
+bfeed.fit <- survfit(Surv(duration, delta) ~ poverty, data = subset(bfeed, duration > 20))
+
+bfeed.poverty.all <- ggsurv(bfeed.fit) +
+  ggtitle('Breast Feeding Duration by Poverty, Duration > 20 Weeks') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+
+grid.arrange(bfeed.poverty.short, bfeed.poverty.all, nrow = 2)
+@
+\end{center}
+\end{frame}
+
+\begin{frame}{Kaplan-Meier Max Likelihood Estimator for $\hat S(t)$}
 \begin{tabular}{ll}
@@
 \begin{eqnarray*}
-  \hat S(t_0) & = & 1 \\
-  \hat S(t)   & = & \prod_{t_i < t} \frac{n_i - d_i}{n_i} \\
-              & = & \frac{n_i - d_i}{n_i} \cdot S(t_{i - 1}) \\ \\
-    n_{i + 1} & = & n_i - d_i - c_i
+                             n_{i + 1} & = & n_i - d_i - c_i \\
+                             \hat S(0) & = & 1 \\
+                             \hat S(t) & = & \prod_{t_i = 0}^{t} \frac{n_i - d_i}{n_i} \\
+                                       & = & \frac{n_i - d_i}{n_i} \cdot S(t_{i - 1}) \\
+  \mathrm{Var}\left[ \hat S(t) \right] & = & \sum_{t_i = 0}^{t} \frac{d_i}{n_i (n_i - d_i)}
 \end{eqnarray*}
@@
 (Cox) Proportional Hazards Model
+% semi-parametric
+% hazards function given covariates (factors)
+% H_0: all \Beta_i = 0
 
@@
 Accelerated Failure Time Model
+% parametric, specifies log(T)
+%   exponential
+%   weibull
+%   lognormal
+%   log-log
+%   gamma
 

### 2014-04-03 presentation.Rnw
Refactoring!

@@
 \documentclass[12pt]{beamer}
+\usepackage{nicefrac}
 
@@
   & $OR(\mbox{no event until } t)$
-  & $\lambda(t)$ & = & $^{f(t)} / _{S(t)}$ \\
+  & $\lambda(t)$ & = & $\nicefrac{f(t)}{S(t)}$ \\
 Cum. Haz.
@@
 <<bfeed-by-education, echo = FALSE, fig = TRUE, height = 4>>=
-bfeed.short$college <- bfeed.short$yschool > 12
-bfeed.fit <- survfit(Surv(duration, delta) ~ college, data = bfeed.short)
-
-ggsurv(bfeed.fit) +
+ggsurv(survfit(Surv(duration, delta) ~ yschool > 12, data = bfeed.short)) +
   ggtitle('Breast Feeding Duration by Education') +
@@
 <<bfeed-by-poverty, echo = FALSE, fig = TRUE, height = 5>>=
-bfeed.fit <- survfit(Surv(duration, delta) ~ poverty, data = bfeed.short)
-
-bfeed.poverty.short <- ggsurv(bfeed.fit) +
+bfeed.poverty.short <-
+  ggsurv(survfit(Surv(duration, delta) ~ poverty, data = bfeed.short)) +
   ggtitle('Breast Feeding Duration by Poverty, Duration < 30 Weeks') +
@@
 
-bfeed.fit <- survfit(Surv(duration, delta) ~ poverty, data = subset(bfeed, duration > 20))
-
-bfeed.poverty.all <- ggsurv(bfeed.fit) +
+bfeed.poverty.all <-
+  ggsurv(survfit(Surv(duration, delta) ~ poverty, data = subset(bfeed, duration > 20))) +
   ggtitle('Breast Feeding Duration by Poverty, Duration > 20 Weeks') +

### 2014-04-03 presentation.Rnw
Finish models
they're really quite simple!
I call this done-done!

@@
 
+\begin{frame}{Nelson-Aalen Estimator for $H(t)$}
+\begin{eqnarray*}
+                             \hat H(0) & = & 0 \\
+                             \hat H(t) & = & \sum_{t_i = 0}^{t} \frac{d_i}{n_i} \\
+                                       & = & \frac{d_i}{n_i} + H(t_{i - 1}) \\
+  \mathrm{Var}\left[ \hat H(t) \right] & = &
+  \sum_{t_i = 0}^{t} \frac{(n_i - d_i) d_i}{(n_i - 1) n_i^2}
+\end{eqnarray*}
+\end{frame}
+
 \begin{frame}{Models}
 (Cox) Proportional Hazards Model
-% semi-parametric
-% hazards function given covariates (factors)
-% H_0: all \Beta_i = 0
-
+% examines cumulative hazards
+% vary lambda_0: different distributions
+% eg. lambda_0 ~ constant => exponential
+% eg. lambda_0 ~ weibull => weibull
 $$\lambda(t | \mathbf{x}) = \lambda_{0}(t) \exp \left( \mathbf{x' \beta} \right)$$
 
+\vspace{0.5in}
+
 Accelerated Failure Time Model
-% parametric, specifies log(T)
-%   exponential
-%   weibull
-%   lognormal
-%   log-log
-%   gamma
-
+% examines survival time
+% vary e: different distributions
+% eg. e_i ~ normal => log-normal
 $$\log T = \mathbf{x' \beta} + \mathbf{\epsilon}$$
-
-Exponential/Weibull Model
-
-$$\lambda(t, \mathbf{x}) = \lambda_{0} \exp \left( \mathbf{x' \beta} \right)$$
 \end{frame}
 
+\begin{frame}[fragile]{Confirmatory Statistics}
+Are all the hazard curves the same?
+
+\begin{eqnarray*}
+  H_0 & : & \text{all } H_i = H_j \\
+  H_a & : & \text{some } H_i \neq H_j
+\end{eqnarray*}
+
+\scriptsize
+<<bfeed-test-correspondence>>=
+survdiff(Surv(duration, delta) ~ poverty, data = bfeed.short)
+@
+\end{frame}
+
+\begin{frame}[fragile]{Confirmatory Statistics}
+Are two survival curves the same?
+
+\begin{eqnarray*}
+  H_0 & : & H_{s \leq 12} = H_{s > 12} \\
+  H_0 & : & H_{s \leq 12} \neq H_{s > 12}
+\end{eqnarray*}
+
+\tiny
+<<bfeed-test-single>>=
+summary(coxph(Surv(duration, delta) ~ yschool > 12, data = bfeed.short))
+@
+\end{frame}
+
+%%% http://data.princeton.edu/wws509/notes/c7.pdf
 %%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf

### 2014-04-03 presentation.Rnw
Oh shit: truncate, don't throw away data!
More fixing derps

@@
 # FDA recommends 6 months of breastfeeding ~ 27 weeks
-bfeed.short <- subset(bfeed, duration < 32)
+bfeed.short <- bfeed
+bfeed.short$duration <- with(bfeed.short, ifelse(duration > 36, 36, duration))
 @
@@
 <<bfeed-test-correspondence>>=
-survdiff(Surv(duration, delta) ~ poverty, data = bfeed.short)
+survdiff(Surv(duration, delta) ~ poverty, data = bfeed)
 @
@@
 <<bfeed-test-single>>=
-summary(coxph(Surv(duration, delta) ~ yschool > 12, data = bfeed.short))
+summary(coxph(Surv(duration, delta) ~ yschool > 12, data = bfeed))
 @

### 2014-04-03 presentation.Rnw
Oh shit more fixing discarded data

@@
 
+bfeed.tail <- bfeed
+bfeed.tail$duration <- with(bfeed.tail, ifelse(duration > 20, duration, 20))
 bfeed.poverty.all <-
-  ggsurv(survfit(Surv(duration, delta) ~ poverty, data = subset(bfeed, duration > 20))) +
+  ggsurv(survfit(Surv(duration, delta) ~ poverty, data = subset(bfeed.tail, duration > 20))) +
   ggtitle('Breast Feeding Duration by Poverty, Duration > 20 Weeks') +

### 2014-04-12 paper.Rnw
Stub paper outline

\documentclass{article}

\title{Survival Analysis}
\author{Brian To}
\date{Spring 2014}

\begin{document}
  \SweaveOpts{concordance=TRUE}
  \maketitle

\section*{Introduction}
% Objective of Paper
% Background on Statistical Technique

\section*{Methodology}
% Overview of Technique
% Application to this data set (variables, parameters, what “practical” questions will you attempt to answer with this analysis)

\section*{Results}
% Exploratory Analysis (“graphical”: graphs, descriptive statistics)
% Statistical Analysis (“analytical”: hypotheses, test statistics, p-values, conclusions)
% Validation of Assumptions (graphs and/or tests)

\section*{Discussion}
% Explain conclusions and interpretations in layman's terms.

%%% References
% List all books, journal articles and web pages used.

%%% Appendices
% Full data set.
% Output from analyses.

\end{document}

### 2014-04-12 paper.Rnw
More outlining
Add references and some intro

@@
-\documentclass{article}
+\documentclass[twocolumn]{article}
 
@@
 \section*{Introduction}
+
+Survival analysis is a relatively recent branch of statistics that tries to answer the question "what proportion of the population will survive until time $t$?" and similar questions.
+For example, "what proportion of Toyota buyers still own and use their car 10 years later?" and "what proportion of Toyota versus Ford buyers still own and use their car 10 years later?" To answer these types of questions, survival analysis focuses on measuring the \emph{duration until event}.
+In the car example, the measurement is the duration of ownership with use and the event is relinquishing ownership or stopping use such as in the case of selling or scraping the car.
+
+Survival analysis has roots in actuarial sciences where life tables were used in calculating insurance rates centuries ago. Only within the last century has Survival Analysis become more popular.
+% Background on Statistical Technique
+% * Tries to answer "what proportion of the population will survive until x?"
+% * started with actuarial tables
+% * WWII - used for reliability analysis
+% * Ed Kaplan & Paul Meier developed S.A. in cancer trials
+%   + This is when S.A. took off, much interest
+%   + Developed Kaplan-Meier survival estimator
+% * Used in other fields (economics, engineering, sociology, ...)
 % Objective of Paper
-% Background on Statistical Technique
+% * Guide one enoguh to be able to:
+%   + design survival studies
+%   + read survival/hazard curves
+%   + estimate kaplan-meier
+%   + use R to perform tests
 
@@
 % Overview of Technique
+% * Concepts
+%   +  Problems
+%   + Inputs
+%   + Censoring
+%     - Staggered entry into study
+%     - Something prevents event (ex. end of study)
+% * Formulas of interest
+%   + f(x), S(x), lambda(x), Lambda(x)
+%   + Kaplan-Meier estimator for S(x)
+%   + Nelson-Aalen estimator for Lambda(x)
+% * Survival models
+%   + accelerated time
+%   + proportional hazards
 % Application to this data set (variables, parameters, what “practical” questions will you attempt to answer with this analysis)
+% * National Labor Survey of Youth
+% * Examined pregnancies
+%   + Only first-born
+%   + Only gestation of 20 to 45 weeks
+%   + Only complete responses
 
@@
 % Exploratory Analysis (“graphical”: graphs, descriptive statistics)
+% * Survival of some college vs no college
+% * Survival by poverty (before 27 weeks, after 27 weeks)
 % Statistical Analysis (“analytical”: hypotheses, test statistics, p-values, conclusions)
@@
 % List all books, journal articles and web pages used.
+%%% http://data.princeton.edu/wws509/notes/c7.pdf
+%%% http://www2.sas.com/proceedings/sugi26/p244-26.pdf
+%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
+%%% http://www.jehps.net/juin2009/Aalenetal.pdf
 

### 2014-04-14 paper.Rnw
Finished background of introduction

@@
 
-Survival analysis has roots in actuarial sciences where life tables were used in calculating insurance rates centuries ago. Only within the last century has Survival Analysis become more popular.
+Survival analysis has roots in actuarial sciences where life tables were used in calculating insurance rates centuries ago.
+However, there has not been much advancement in the field until the last century.
+In World War II, military technology prowess did not keep up with its reliability until survival analysis was used to increase the reliability of equipment.
+Advances, however, were only limited to uses in the military and not embraced by the public.
+In the mid 1970s, Major Greenwood published a paper on cancer incidence and mortality by various types.
+Edward Kaplan and Paul Meier took notice of this study and developed what is known as the Kaplan-Meier estimator for the survival rate.
+After Kaplan and Meier published their landmark survival analysis paper, interest in the subject exploded and spread to other fields of study besides medical and engineering such as sociology and economics.
 % Background on Statistical Technique
@@
 %%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3227332/
+%%% http://www.ncbi.nlm.nih.gov/pmc/articles/PMC3059453/
 %%% http://www.jehps.net/juin2009/Aalenetal.pdf
+%%% www.worldofstatistics.org/files/2012/12/Stat2013.pptx
+%%% http://books.google.com/books?id=1iI5rHtCT-4C&lpg=PA112&ots=2jMXda6Zjy&dq=%22major%20greenwood%22%20cancer&pg=PA112#v=onepage&q=%22major%20greenwood%22%20cancer&f=false
 

### 2014-04-14 paper.Rnw
Start on methodology.

@@
 \documentclass[twocolumn]{article}
+\usepackage[margin=1in]{geometry}
 
@@
 After Kaplan and Meier published their landmark survival analysis paper, interest in the subject exploded and spread to other fields of study besides medical and engineering such as sociology and economics.
+
+The objective of this paper is to provide an introduction to survival analysis, including the design of survival studies, estimation and interpretation of survival and hazard curves, and using R to perform significance tests.
 % Background on Statistical Technique
@@
 \section*{Methodology}
+
+\subection*{Censoring}
+Data collection steps in experiments may not yield perfect data.
+For instance, data for certain subjects may be missing or the subject may drop out in the middle of the study.
+These types of data imperfections are called "censoring" in survival analysis.
+There are
 % Overview of Technique

### 2014-04-15 paper.Rnw
Add censoring section

@@
 
-\subection*{Censoring}
+\subsection*{Censoring}
 Data collection steps in experiments may not yield perfect data.
@@
 These types of data imperfections are called "censoring" in survival analysis.
-There are
+There are three types of censoring: left, right, and interval.
+The difference between these types of censoring is based on when the censoring occurs.
+
+Left censoring is when the subject has the event before the start of the study.
+For example, a study that looks at when teenagers have their first alcoholic drink after entering high school may have left-censored data if there are teenagers that have already started drinking in middle school.
+Left censoring typically happens when the event is hard to detect prior to the study.
+
+Right censoring is when the subject has the event after the study ends.
+For example, a study examining cancer survival may have right censored data if subjects survive past the end of the study.
+Right censoring is more common than left censoring and typically occurs in longitudinal studies or studies that allow for staggered entry to the study.
+
+Interval censoring is when the subject has the event within a time interval and the true time of event is not known.
+For example, a study examining the onset of sexually transmitted diseases for sex workers in third world countries may have interval censored data because determining the exact time of acquisition is impossible and acquisition can only be determined by check-ins which does not cause acquisition of STDs.
 % Overview of Technique

### 2014-04-24 paper.Rnw
Add formulas

@@
 \documentclass[twocolumn]{article}
+
 \usepackage[margin=1in]{geometry}
+\usepackage{nicefrac}
 
@@
 For example, a study examining the onset of sexually transmitted diseases for sex workers in third world countries may have interval censored data because determining the exact time of acquisition is impossible and acquisition can only be determined by check-ins which does not cause acquisition of STDs.
+
+\subsection*{Concepts}
+Assume that $f(t)$ represents the probability of the event occurring at time $t$.
+Each $f(t)$ is independent of other $f(t)$ values.
+For example, $f(2)$ has no relation to $f(3)$.
+$f$'s shape may be defined some concrete function, but future values may not be computed from previous values, such as $f(t) = 0.7 f(t - 1)$.
+
+The survival function $S(t)$ denotes the probability that a subject survives from $t_0$ to $t$.
+For a subject to survive in this interval, it must not have an event in any subinterval $t_0$, $t_1$, $t_2$, \ldots, $t$.
+Since the probability of not having an event in a interval is $1 - f(t)$, we can represent $S(t)$ as
+
+\[ S(t) = \prod_{t_i = 0}^{t} 1 - f(t) \]
+
+More intuitively, we can say that the probability of surviving $t$ is the probability of surviving until $t$ and surviving $t$ itself. In other words
+
+\[ S(t) = S(t - 1) \cdot (1 - f(t)) \]
+
+The hazard function $\lambda(t)$ represents the odds of surviving until time $t$ and then experiencing an event in time $t$.
+This probability can be written as $P(\mbox{event} | \mbox{survive until } t)$.
+We may write this conditional probability as
+
+\[ \lambda(t) = \nicefrac{f(t)}{S(t)} \]
+
+The cumulative hazard function $\Lambda(t)$ represents the odds of experiencing an event in the interval until $t$.
+In other words, it is th cumulative risk of the event until $t$.
+A way to express this is the sum of risk for every interval until $t$.
+More succintly,
+
+\[ \Lambda(t) = \sum_{t_i = 0}^{t} \lambda(t) \]
+
+It is important to note that $\lambda$ and $\Lambda$ are not probabilities, but \emph{odds ratios}. The range of these functions are not $[0, 1]$ but $[0, \infty)$.
+
+\subsection*{Estimators}
+
 % Overview of Technique

### 2014-04-24 paper.Rnw
Add Kaplan Meier, start working on graphs.

@@
 \usepackage{nicefrac}
+\usepackage{tabularx}
 
@@
 \maketitle
+<<echo = FALSE>>=
+library(ggplot2)
+library(gridExtra)
+ws <- read.csv('worksheet.csv', header = TRUE)
+ws$id <- sprintf('%02d', 1:nrow(ws))
+@
 
@@
 
+Edward Kaplan and Paul Meier's landmark paper on the Kaplan-Meier estimator drove survival analysis forward since it was the first accepted estimator for $S(t)$ that also handles censored data.
+Recall that $S(t) = S(t - 1) \cdot (1 - f(t))$.
+$f(t)$ in Kaplan-Meier can be expressed as the number of events divided by the number of subjects at risk at time $t$.
+In other words,
+
+\[ f(t_i) =  \frac{d_i}{n_i} \]
+
+where $d_i$ represents the number of events at $t$ and $n_i$ represents the number of subjects at risk at $t$.
+But what happens if there are censored data points for time $t$?
+Since there is no actual event, it would not make sense for those points to affect the current survival.
+In the next interval, the number of subjects at risk $n_{i + 1}$ should be affected because those censored subjects are no longer available at $t_{i + 1}$.
+If we denote this quantity of censored data points as $c_i$, we may say that the number of subjects at risk in the next interval is the same as this interval, but without those that experienced the event and those that are censored.
+In other words,
+
+\[ n_{i + 1} = n_i - d_i - c_i \]
+
+While many software packages can easily create survival estimates, computing $\hat S(t)$ can be relatively straightforward using tables. Let's examine sample data shown as Figure \ref{fig:ws}.
+
+\begin{figure}[p]
+<<ws, echo = FALSE, fig = TRUE, height = 3>>=
+ggplot(ws, aes(y = id, yend = id, colour = event)) +
+  ggtitle('Event Duration') +
+  xlab('duration') +
+  ylab('survival') +
+  geom_segment(aes(
+    x = 0,
+    xend = duration,
+    order = id)) +
+  geom_point(aes(
+    x = duration), size = 2)
+@
+\caption{thing}
+\label{fig:ws}
+\end{figure}
+
 % Overview of Technique

### 2014-04-27 paper.Rnw
Tweak image and include algorithm steps

@@
 \usepackage{tabularx}
+\usepackage{graphicx}
 
@@
 
-While many software packages can easily create survival estimates, computing $\hat S(t)$ can be relatively straightforward using tables. Let's examine sample data shown as Figure \ref{fig:ws}.
+While many software packages can easily create survival estimates, computing $\hat S(t)$ can be straightforward using tables. Let's examine sample data shown below.
 
-\begin{figure}[p]
-<<ws, echo = FALSE, fig = TRUE, height = 3>>=
+<<ws, echo = FALSE, fig = TRUE, include = FALSE>>=
 ggplot(ws, aes(y = id, yend = id, colour = event)) +
+  theme(legend.position = "none") +
   ggtitle('Event Duration') +
@@
   ylab('survival') +
+  guides(fill = FALSE) +
   geom_segment(aes(
@@
 @
-\caption{thing}
+\begin{figure}[h]
+\begin{center}
+\includegraphics[keepaspectratio=true, width=3in]{paper-ws.pdf}
+\end{center}
+\caption{Sample Survival Data}
 \label{fig:ws}
 \end{figure}
+
+In this sample data, teal endpoints represent an event observation and red endpoints repreesnt a censored observation.
+We begin by creating a table with the following quantities: $n_i$, $d_i$, $c_i$, $\nicefrac{d_i}{n_i}$, $\nicefrac{n_i - d_i}{n_i}$, and $\hat S(t_i)$.
+Afterwards, we may perform the following algorithm:
+
+\begin{enumerate}
+  \item \emph{Calculate $n_i$.} $n_i$ is the total number of subjects that are at risk in this interval.
+If a vertical line is drawn at $t_i$, $n_i$ is the total number of duration lines the vertical line touches.
+This includes subjects that either have an event or are censored in this interval.
+Equivalently, you may also calculate $n_i$ by taking $n_{i - 1}$ and subtracting the number of events and censors in the previous duration.
+  \item \emph{Calculate $d_i$.} $d_i$ is the number of teal duration endings at $t_i$.
+  \item \emph{Calculate $c_i$.} $c_i$ is the number of red duration endings at $t_i$.
+  \item \emph{Calculate $\nicefrac{d_i}{n_i}$ and $\nicefrac{n_i - d_i}{d_i}$} where the former represents the failure rate $f(t_i)$ and the latter represents the probability of not failing at that specific interval.
+Only the latter is used to calculate $\hat S(t_i)$.
+  \item \emph{Calculate $S(t_i)$.} Recall that $S(t_i) = \nicefrac{n_i - d_i}{d_i} \cdot S(t_{i - 1})$.
+If $S(t_{i - 1})$ is not defined, use 1 indicating that everyone survived the imaginary duration.
+  \item \emph{Repeat until $n_i$ is zero.}
+\end{enumerate}
 

### 2014-04-27 paper.Rnw
Add sample computation table

@@
 library(gridExtra)
+library(survival)
+library(xtable)
 ws <- read.csv('worksheet.csv', header = TRUE)
@@
 
+Below is the result of this algorithm on the sample.
+
+<<ws-table, echo = FALSE, results = tex>>=
+ws.fit <- survfit(Surv(duration, event) ~ 1, data = ws)
+
+ws.table <- with(ws.fit, data.frame(
+  n = n.risk,
+  d = n.event,
+  c = n.censor,
+  f = n.event / n.risk,
+  nf = 1 - (n.event / n.risk),
+  s = surv))
+
+colnames(ws.table) <-
+  c("$n_i$", "$d_i$", "$c_i$", "$\\nicefrac{d}{n}$", "$\\frac{n-d}{n}$", "$\\hat S_i$")
+
+print(xtable(ws.table),
+  sanitize.text.function = function(x) {
+    return(x)
+  })
+@
+
 % Overview of Technique

### 2014-04-27 paper.Rnw
Added blurb for nelson-aalen

@@
 
+The values of $\hat S_i$ in the table give the estimates for the Kaplan-Meier survival estimator.
+
+Similarly, the cumulative risk estimator $\hat \Lambda{t}$ can be estimated given our sample data.
+An estimate for the cumulative risk is the cumulative sum of $\frac{n - d}{n}$ which can be easily computed from the table above.
+This is called the Nelson-Aalen estimator.
 % Overview of Technique

### 2014-04-28 paper.Rnw
Added data set section

@@
 library(survival)
+library(KMsurv)
 library(xtable)
+
 ws <- read.csv('worksheet.csv', header = TRUE)
 ws$id <- sprintf('%02d', 1:nrow(ws))
+
+data(bfeed)
 @
@@
 This is called the Nelson-Aalen estimator.
+
+\subsection*{Models}
+
+TODO
 % Overview of Technique
@@
 %   + Only complete responses
+
+\section*{Data Set}
+
+We will examine data from the National Labor Survey of Youth.
+The data set looks at the relation between the duration of breast feeding and various factors.
+This data set examines pregnancies before 1983 and includes only complete data points (no data missing).
+According to the original researchers, the year restriction is put in place to mitigate ``recall problems.'' Also, this data is restricted to first-born children and to children whose gestation is between 20 and 45 weeks.
+With these restrictions, there are a total of 927 data points.
+
+Breastfeeding duration is measured in weeks.
+The study tracks the following factors: race (white, black, other), poverty, somking, drinks alcohol, age of mother, mother's year of birth, mother's education level, and if the mother has prenatal car after the 3rd month of pregnancy.
+
+Race is denoted as the integers 1 to 3 where 1 represents white, 2 represents black, and 3 represents other.
+Poverty, smoking, drinks alcohol, and prenatal care are denoted as booleans where 1 represents a ``yes'' to the category and 0 represents ``no''.
+For example, a 1 in the smoking category means the mother is a smoker.
+Age of mother is in years.
+The mother's year of birth is represented as the last two digits of the year she was born.
+For example, a value of ``84'' indicates that the mother was born in 1984.
+Education level is represented as the number of completed years of education starting from 1st grade through college, if applicable.
+
+A sample of the full data set is shown below.
+
+<<bfeed-raw, echo = FALSE, results = tex>>=
+bfeed.show <- bfeed[c(556, 108, 21, 513, 109, 16), ]
+rownames(bfeed.show) <- 1:nrow(bfeed.show)
+bfeed.show <- t(bfeed.show)
+
+xtable(bfeed.show)
+@
 

### 2014-04-28 paper.Rnw
Added visualization of bfeed survival

@@
 library(survival)
+library(GGally)
 library(KMsurv)
@@
 data(bfeed)
+
+bfeed.short <- bfeed
+bfeed.short$duration <- with(bfeed.short, ifelse(duration > 36, 36, duration))
 @
@@
 Breastfeeding duration is measured in weeks.
+An event (delta) is defined as full weaning of the infant from breast milk.
 The study tracks the following factors: race (white, black, other), poverty, somking, drinks alcohol, age of mother, mother's year of birth, mother's education level, and if the mother has prenatal car after the 3rd month of pregnancy.
@@
 Education level is represented as the number of completed years of education starting from 1st grade through college, if applicable.
+
+The range of durations for breastfeeding is from 1 week to 192 weeks.
+Only 12\% of subjects had breastfeeding durations greater than 36 weeks.
+If we were to graph $\hat S(t)$, 20\% of the x-axis will contain 88\% of the data, which makes detecting patterns in the data visually very difficult.
+For this reason, durations greater than 36 weeks will be treated as 36 weeks.
 
@@
 \section*{Results}
+
+Below is a graph of survival data for all points with no explanatory variables in the model.
+
+<<bfeed-survival, echo = FALSE, fig = TRUE, include = FALSE>>=
+bfeed.fit <- survfit(Surv(duration, delta) ~ 1, data = bfeed.short)
+
+ggsurv(bfeed.fit) +
+  ggtitle('Breast Feeding Duration') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+@
+\begin{figure}[h]
+\begin{center}
+\includegraphics[keepaspectratio=true, width=3in]{paper-bfeed-survival.pdf}
+\end{center}
+\caption{Breast Feeding Survival Data}
+\label{fig:bfeed-survival}
+\end{figure}
+
+The red dashed line represents the Food and Drug Administration's recommended duration of breastfeeding, 6 months.
+The steep drop off at 36 weeks is due to the data clipping used to make the graphs easier to interpret.
+The red crosses indicate censored observations and the dashed black line is the confidence band of the Kaplan-Meier estimator.
+
 % Exploratory Analysis (“graphical”: graphs, descriptive statistics)

### 2014-04-28 paper.Rnw
Start analysis.

@@
 
-\section*{Results}
+\section*{Results and Discussion}
 
@@
 
+Notice that the survival rate drops off sharply after the first feel weeks.
+For weeks 0 to 5, there is a 25\% drop in survival.
+In contrast, there is a similar drop in survival for weeks 10 to 20.
+Casually, we can say that the survival rate for breast feeding improves for the mothers who breast feed longer.
+Likewise, we can say that the expected duration of breastfeeding is around 10 weeks and that less than 25\% of mothers breast feed for at least the FDA recommended breastfeeding duration.
+
 % Exploratory Analysis (“graphical”: graphs, descriptive statistics)
@@
 % Validation of Assumptions (graphs and/or tests)
-
-\section*{Discussion}
-% Explain conclusions and interpretations in layman's terms.
 

### 2014-04-28 paper.Rnw
Add bfeed education and some analysis

@@
 
+One question we may ask is ``Do college-educated women breastfeed longer?'' Intuitively, we may reason that college-educated women are more aware of parenting issues and have more access to parenting resources, so they are more likely to know to breastfeed for a long duration and have the support to breastfeed for a long duration.
+Or perhaps it is the reverse since less-educated women may depend more on parents for parenting support and their parents may provide more support for breastfeeding long durations.
+
+<<bfeed-education, echo = FALSE, fig = TRUE, include = FALSE>>=
+ggsurv(survfit(Surv(duration, delta) ~ yschool > 12, data = bfeed.short)) +
+  ggtitle('Breast Feeding Duration by Education') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+@
+
+The red line represents the survival rate for women with a high school diploma or less and the blue line represents the survival rate for women with some college education.
+
+In this instance, women with some college education tend to breastfeed children longer than women with a high school education or less.
+That is \emph{not} to say \emph{every} woman with a high school diploma or less breast feeds less than \emph{every} woman with some college education.
+Nor can we say that having some college education causes longer breastfeeding durations.
+
+% TODO test
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[keepaspectratio=true, width=3in]{paper-bfeed-education.pdf}
+\end{center}
+\caption{Breast Feeding Survival By Education}
+\label{fig:bfeed-education}
+\end{figure}
+
 % Exploratory Analysis (“graphical”: graphs, descriptive statistics)

### 2014-04-28 paper.Rnw
Add hypothesis test for education

@@
 
-% TODO test
+We may verify this casual relation by performing a log-rank test assuming $\alpha = 0.05$ and where the hypotheses are:
+
+\[ H_0: \Lambda_{ys \leq 12} = \Lambda_{ys > 12} \]
+\[ H_a: \Lambda_{ys \leq 12} \neq \Lambda_{ys > 12} \]
+
+\scriptsize
+<<bfeed-education-test, echo = FALSE>>=
+survdiff(Surv(duration, delta) ~ yschool > 12, data = bfeed.short)
+@
+\normalsize
+
+The $p$ value from the log-rank test is lower than our $\alpha$, which means we can reject $H_0$ that having some college education makes a difference in breastfeeding durations and support our casual conclusion that having some college education indicates longer breastfeeding durations.
 

### 2014-04-29 paper.Rnw
Add poverty analysis

@@
 @
+\begin{figure}[h]
+\begin{center}
+\includegraphics[keepaspectratio=true, width=3in]{paper-bfeed-education.pdf}
+\end{center}
+\caption{Breast Feeding Survival By Education}
+\label{fig:bfeed-education}
+\end{figure}
 
@@
 
+Another questions we may ask is ``Does poverty influence a mother's breastfeeding duration?'' Intuitively, we may guess that more affluent women can afford to breastfeed longer because they can afford maternity leave which allows time for breastfeeding.
+On the other hand, poorer women may breastfeed longer since formula is rather expensive, breastfeeding may be the cheaper alternative.
+
+<<bfeed-poverty, echo = FALSE, fig = TRUE, include = FALSE>>=
+ggsurv(survfit(Surv(duration, delta) ~ poverty, data = bfeed.short)) +
+  ggtitle('Breast Feeding Duration by Poverty, Duration < 36 Weeks') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+@
 \begin{figure}[h]
 \begin{center}
-\includegraphics[keepaspectratio=true, width=3in]{paper-bfeed-education.pdf}
+\includegraphics[keepaspectratio=true, width=3in]{paper-bfeed-poverty.pdf}
 \end{center}
-\caption{Breast Feeding Survival By Education}
-\label{fig:bfeed-education}
+\caption{Breast Feeding Survival By Poverty}
+\label{fig:bfeed-poverty}
 \end{figure}
+
+We can see that the two survival curves for the mothers in poverty and the mothers not in poverty cross many times and do not show any clear distinguishing pattern from each other.
+Casually, we can say that these two survival curves are not different.
+
+However, if we compress not the right tail but the first 20 weeks of breastfeeding, we arrive at a different conclusion.
+
+<<bfeed-poverty-tail, echo = FALSE, fig = TRUE, include = FALSE>>=
+bfeed.tail <- bfeed
+bfeed.tail$duration <- with(bfeed.tail, ifelse(duration > 20, duration, 20))
+
+ggsurv(survfit(Surv(duration, delta) ~ poverty, data = subset(bfeed.tail, duration > 20))) +
+  ggtitle('Breast Feeding Duration by Poverty, Duration > 20 Weeks') +
+  xlab('Duration in Weeks') +
+  ylab('Survival') +
+  geom_vline(aes(xintercept = 27),
+    colour = 'red',
+    linetype = 'longdash')
+@
+\begin{figure}[h]
+\begin{center}
+\includegraphics[keepaspectratio=true, width=3in]{paper-bfeed-poverty-tail.pdf}
+\end{center}
+\caption{Breast Feeding Survival By Poverty}
+\label{fig:bfeed-poverty-tail}
+\end{figure}
+
+In figure \ref{fig:bfeed-poverty-tail}, we see that mothers who are in poverty generally breastfeed longer than those who are not in poverty.
+From this finding, we can casually claim that poverty increases breastfeeding survival rates only after some threshold duration.
+
+If we perform a log-rank test to examine if the survival functions are different, we find that there is little evidence to reject that the survival curves are the same.
+We also reach the same conclusion if we truncate either the right tail or left tail.
+Even though the right tail hints that mothers in poverty breastfeed for longer, there is insufficient evidence to suggest that it is the case.
+
+\[ H_0: \Lambda_{p = 1} = \Lambda_{p = 0} \]
+\[ H_a: \Lambda_{p = 1} \neq \Lambda_{p = 0} \]
+
+\scriptsize
+<<bfeed-poverty-test, echo = FALSE>>=
+survdiff(Surv(duration, delta) ~ poverty, data = bfeed)
+survdiff(Surv(duration, delta) ~ poverty, data = bfeed.short)
+survdiff(Surv(duration, delta) ~ poverty, data = bfeed.tail)
+@
+\normalsize
 

### 2014-04-29 presentation.Rnw
Add missing definition for c_i
Add poverty analysis

@@
   ggsurv(survfit(Surv(duration, delta) ~ poverty, data = bfeed.short)) +
-  ggtitle('Breast Feeding Duration by Poverty, Duration < 30 Weeks') +
+  ggtitle('Breast Feeding Duration by Poverty, Duration < 36 Weeks') +
   xlab('Duration in Weeks') +
@@
   $d_i$ & \# events in $t_i$ \\
+  $c_i$ & \# censored subjects in $t_i$ \\
   $n_i$ & \# subjects \emph{at risk} in $t_i$ \\

### 2014-05-02 paper.Rnw
More stuff on models

@@
 
-TODO
+Survival models can generally be divided into two types: proportional hazards and accelerated failure time models.
+
+Proportional hazards models model the effect of explanatory variables on the hazard rate $\lambda$.
+
+\[ \lambda(t | \mathbf{x}) = \lambda_{0}(t) \exp \left( \mathbf{x' \beta} \right) \]
+
+In this model, $\mathbf{x}$ is our matrix of factor values and $\mathbf{\beta}$ is our matrix of factor coefficients.
+$\lambda_{0}(t)$ is our baseline hazard rate, or the hazard rate we would expect if the factors supplied have no effect on $\lambda(t | \mathbf{x})$.
+
+We can view the $\exp \left( \mathbf{x' \beta} \right)$ factor in the model above as a modifier for the baseline hazard rate $\lambda_{0}(t)$ that streatches and shrinks the baseline hazard vertically.
+In expanded form, this exponential term would look like $\exp \left( x_1 \beta_1 + x_2 \beta_2 + x_3 \beta_3 + \ldots \right)$ which would simplify to $\exp(x_1 \beta_1) \exp(x_2 \beta_2) \exp(x_3 \beta_3) \ldots$.
+In this simplified form, we can see that the scaling factors are not additive, but multiplicative.
+Also, if an $x\beta$ were to be 0, the resulting scaling factor would be $\exp(0) = 1$ which is the same as saying ``this factor has no effect on the hazard rate.''
+
+Variants of the proportional hazards model can be made by varying the baseline hazard's distribution.
+For example, if $\lambda_{0}$ follows an exponential distribution, then the hazard function $\lambda$ follows a Weibull distribution.
+Likewise, if $\lambda_{0}$ is constant, then $\lambda$ follows an exponential distribution.
+
+Accelerated failure time models model the effect of explanaory variables on the time-to-event $T$.
+
+\[ \log T = \mathbf{x' \beta} + \mathbf{\epsilon} \]
+
+Like the proportional hazards model, $\mathbf{x}$ is our matrix of factor values and $\mathbf{\beta}$ is our matrix of factor coefficients.
+$\epsilon$ represents error in our model, similar to the error expressed in linear models.
+
+We can view the $\mathbf{x' \beta}$ factor in the model above as a modifier for the time-to-event $T$ that streatches and shrinks the survival rate horizontally.
+
 % Overview of Technique

### 2014-05-02 paper.Rnw
Copy-paste-modify section for accel. fail. time model.

@@
 We can view the $\mathbf{x' \beta}$ factor in the model above as a modifier for the time-to-event $T$ that streatches and shrinks the survival rate horizontally.
+In expanded form, this linear term would look like $x_1 \beta_1 + x_2 \beta_2 + x_3 \beta_3 + \ldots$ which would simplify to $\exp(x_1 \beta_1) \exp(x_2 \beta_2) \exp(x_3 \beta_3) \ldots$ when we exponentiate both sides, ignoring the error term.
+In this form, we can see that the scaling factors are not additive, but multiplicative.
+If an $x\beta$ were to be 0, the resulting scaling factor would be $\exp(0) = 1$ which is the same as saying ``this factor has no effect on the time-to-event.''
 

### 2014-05-08 paper.Rnw
Finish off models section
Some changes

@@
 For example, "what proportion of Toyota buyers still own and use their car 10 years later?" and "what proportion of Toyota versus Ford buyers still own and use their car 10 years later?" To answer these types of questions, survival analysis focuses on measuring the \emph{duration until event}.
-In the car example, the measurement is the duration of ownership with use and the event is relinquishing ownership or stopping use such as in the case of selling or scraping the car.
+In the car example, the measurement is the duration of ownership with use and the event is relinquishing ownership or stopping use such as in the case of selling or scrapping the car.
 
@@
 In World War II, military technology prowess did not keep up with its reliability until survival analysis was used to increase the reliability of equipment.
-Advances, however, were only limited to uses in the military and not embraced by the public.
-In the mid 1970s, Major Greenwood published a paper on cancer incidence and mortality by various types.
+Advances, however, were limited to uses in the military and not embraced by the public.
+In the mid 1970s, Major Greenwood published a paper on cancer incidence and mortality by various types such as breast, uterus, rectum, tongue and mouth.
 Edward Kaplan and Paul Meier took notice of this study and developed what is known as the Kaplan-Meier estimator for the survival rate.
-After Kaplan and Meier published their landmark survival analysis paper, interest in the subject exploded and spread to other fields of study besides medical and engineering such as sociology and economics.
+After Kaplan and Meier published their landmark survival analysis paper, interest in the subject exploded and spread to other fields of study in addition to medical and engineering such as sociology and economics.
 
@@
 For example, $f(2)$ has no relation to $f(3)$.
-$f$'s shape may be defined some concrete function, but future values may not be computed from previous values, such as $f(t) = 0.7 f(t - 1)$.
+$f$'s shape may be defined by some concrete function, but future values may not be computed from previous values, such as $f(t) = 0.7 f(t - 1)$.
 
-The survival function $S(t)$ denotes the probability that a subject survives from $t_0$ to $t$.
-For a subject to survive in this interval, it must not have an event in any subinterval $t_0$, $t_1$, $t_2$, \ldots, $t$.
+The survival function $S(t)$ denotes the probability that a subject survives from $t_0$ to $t$ where $t_0$ is the start of the study and $t$ is the current time.
+For a subject to survive to the end of this interval, it must not have an event in any subinterval $t_0$, $t_1$, $t_2$, \ldots, $t$.
 Since the probability of not having an event in a interval is $1 - f(t)$, we can represent $S(t)$ as
 
-\[ S(t) = \prod_{t_i = 0}^{t} 1 - f(t) \]
+\[ S(t) = \prod_{t_i = 0}^{t} \left( 1 - f(t) \right) \]
 
-More intuitively, we can say that the probability of surviving $t$ is the probability of surviving until $t$ and surviving $t$ itself. In other words
+More intuitively, we can say that the probability of surviving past time $t$ is the probability of surviving until $t$ and surviving $t$ itself. In other words
 
@@
 
-The hazard function $\lambda(t)$ represents the odds of surviving until time $t$ and then experiencing an event in time $t$.
-This probability can be written as $P(\mbox{event} | \mbox{survive until } t)$.
+The hazard function $\lambda(t)$ represents the event rate given survival until time $t$ and then experiencing an event in time $t$.
+This probability can be written as $P(\mbox{event} | \mbox{survive up to } t)$.
 We may write this conditional probability as
@@
 
-The cumulative hazard function $\Lambda(t)$ represents the odds of experiencing an event in the interval until $t$.
+The cumulative hazard function $\Lambda(t)$ represents the event rate of experiencing an event in the interval until $t$.
 In other words, it is th cumulative risk of the event until $t$.
@@
 
-It is important to note that $\lambda$ and $\Lambda$ are not probabilities, but \emph{odds ratios}. The range of these functions are not $[0, 1]$ but $[0, \infty)$.
+It is important to note that $\lambda$ and $\Lambda$ are not probabilities, but \emph{event rates}. The range of these functions are not $[0, 1]$ but $[0, \infty)$.
 
@@
 Similarly, the cumulative risk estimator $\hat \Lambda{t}$ can be estimated given our sample data.
-An estimate for the cumulative risk is the cumulative sum of $\frac{n - d}{n}$ which can be easily computed from the table above.
+An estimate for the cumulative risk is the cumulative sum of $\frac{d}{n}$ which can be easily computed from the table above.
 This is called the Nelson-Aalen estimator.
@@
 
+Variants of the accelerated failure time model can be made by varying the error term's distribution.
+For example, if $\epsilon$ follows a normal distribution, $T$ follows a log-normal distribution.
+If $\epsilon$ follows a logistic distribution, $T$ follows a log-log distribution.
+
 % Overview of Technique
@@
 An event (delta) is defined as full weaning of the infant from breast milk.
-The study tracks the following factors: race (white, black, other), poverty, somking, drinks alcohol, age of mother, mother's year of birth, mother's education level, and if the mother has prenatal car after the 3rd month of pregnancy.
+The study tracks the following factors: race (white, black, other), poverty, smoking, drinks alcohol, age of mother, mother's year of birth, mother's education level, and if the mother has prenatal care after the 3rd month of pregnancy.
 
-Race is denoted as the integers 1 to 3 where 1 represents white, 2 represents black, and 3 represents other.
+Race is coded as the integers 1 to 3 where 1 represents white, 2 represents black, and 3 represents other.
 Poverty, smoking, drinks alcohol, and prenatal care are denoted as booleans where 1 represents a ``yes'' to the category and 0 represents ``no''.
@@
 If we were to graph $\hat S(t)$, 20\% of the x-axis will contain 88\% of the data, which makes detecting patterns in the data visually very difficult.
-For this reason, durations greater than 36 weeks will be treated as 36 weeks.
+For this reason, durations greater than 36 weeks will be censored as 36 weeks.
 
-A sample of the full data set is shown below.
+A portion of the full data set is shown below.
 
@@
 
-Below is a graph of survival data for all points with no explanatory variables in the model.
+Figure \ref{fig:bfeed-survival} is a graph of survival data for all points with no explanatory variables in the model.
 
@@
 
-The red dashed line represents the Food and Drug Administration's recommended duration of breastfeeding, 6 months.
-The steep drop off at 36 weeks is due to the data clipping used to make the graphs easier to interpret.
-The red crosses indicate censored observations and the dashed black line is the confidence band of the Kaplan-Meier estimator.
+The vertical red dashed line represents the Food and Drug Administration's recommended duration of breastfeeding, 6 months.
+The steep drop off at 36 weeks is due to the censoring used to make the graphs easier to interpret.
+The red crosses indicate censored observations and the dashed black line is the 95\% confidence band of the Kaplan-Meier estimator.
 
-Notice that the survival rate drops off sharply after the first feel weeks.
+Notice that the survival rate drops off sharply after the first few weeks.
 For weeks 0 to 5, there is a 25\% drop in survival.
@@
 
-We may verify this casual relation by performing a log-rank test assuming $\alpha = 0.05$ and where the hypotheses are:
+We may verify this casual relation by performing a log-rank test assuming $\alpha = 0.05$. The log-rank test compares hazard rates rather than survival rates:
 
@@
 % List all books, journal articles and web pages used.
+%%% http://www.cabdirect.org/abstracts/19272700028.html
+%%% http://data.princeton.edu/pop509/ParametricSurvival.pdf
 %%% http://data.princeton.edu/wws509/notes/c7.pdf
@@
 %%% www.worldofstatistics.org/files/2012/12/Stat2013.pptx
+%%% http://books.google.com/books?id=xttbn0a-QR8C&lpg=PA15&ots=aZ4QAk7z6f&dq=interpretation%20of%20failure%20rate%20survival%20analysis&pg=PA14#v=onepage&q=interpretation%20of%20failure%20rate%20survival%20analysis&f=false
 %%% http://books.google.com/books?id=1iI5rHtCT-4C&lpg=PA112&ots=2jMXda6Zjy&dq=%22major%20greenwood%22%20cancer&pg=PA112#v=onepage&q=%22major%20greenwood%22%20cancer&f=false

### 2014-05-08 paper.Rnw
Make fixes

@@
 
-Variants of the proportional hazards model can be made by varying the baseline hazard's distribution.
+Variants of the proportional hazards model can be made by varying the distribution of events $f(t)$ or the baseline hazard's distribution $\lambda_0(t)$.
 For example, if $\lambda_{0}$ follows an exponential distribution, then the hazard function $\lambda$ follows a Weibull distribution.
 Likewise, if $\lambda_{0}$ is constant, then $\lambda$ follows an exponential distribution.
+We may define $f(t)$ instead of $\lambda_{0}(t)$ as well because we can derive $\lambda_{0}(t)$ from $f(t)$.
+Choosing which function to define, the event distribution or baseline hazard rate, is a matter of convenience and preference.
 
@@
 If we were to graph $\hat S(t)$, 20\% of the x-axis will contain 88\% of the data, which makes detecting patterns in the data visually very difficult.
-For this reason, durations greater than 36 weeks will be censored as 36 weeks.
+For this reason, durations greater than 36 weeks will be treated as 36 weeks.
 
@@
 The vertical red dashed line represents the Food and Drug Administration's recommended duration of breastfeeding, 6 months.
-The steep drop off at 36 weeks is due to the censoring used to make the graphs easier to interpret.
+The steep drop off at 36 weeks is due to the truncation used to make the graphs easier to interpret.
 The red crosses indicate censored observations and the dashed black line is the 95\% confidence band of the Kaplan-Meier estimator.
@@
 
-We may verify this casual relation by performing a log-rank test assuming $\alpha = 0.05$. The log-rank test compares hazard rates rather than survival rates:
+We may verify this casual relation by performing a log-rank test assuming $\alpha = 0.05$. The hypotheses for this test are:
 
-\[ H_0: \Lambda_{ys \leq 12} = \Lambda_{ys > 12} \]
-\[ H_a: \Lambda_{ys \leq 12} \neq \Lambda_{ys > 12} \]
+\[ H_0: \S_{ys \leq 12} = \S_{ys > 12} \]
+\[ H_a: \S_{ys \leq 12} \neq \S_{ys > 12} \]
 

### 2014-05-08 paper.Rnw
More fixing things
Make tests as figures with captions.

@@
 \usepackage{graphicx}
+\usepackage{cite}
+\usepackage{hyperref}
 
@@
 
-\[ H_0: \S_{ys \leq 12} = \S_{ys > 12} \]
-\[ H_a: \S_{ys \leq 12} \neq \S_{ys > 12} \]
+\[ H_0: S_{ys \leq 12} = S_{ys > 12} \]
+\[ H_a: S_{ys \leq 12} \neq S_{ys > 12} \]
 
@@
 
-\[ H_0: \Lambda_{p = 1} = \Lambda_{p = 0} \]
-\[ H_a: \Lambda_{p = 1} \neq \Lambda_{p = 0} \]
+\[ H_0: S_{p = 1} = S_{p = 0} \]
+\[ H_a: S_{p = 1} \neq S_{p = 0} \]
 
+\begin{figure}[h]
 \scriptsize
-<<bfeed-poverty-test, echo = FALSE>>=
+<<bfeed-poverty-test-normal, echo = FALSE>>=
 survdiff(Surv(duration, delta) ~ poverty, data = bfeed)
+@
+\normalsize
+\label{fig:bfeed-poverty-test-normal}
+\caption{Testing $H_0: S_{p = 1} = S_{p = 0}$ for all data}
+\end{figure}
+
+\begin{figure}[h]
+\scriptsize
+<<bfeed-poverty-test-short, echo = FALSE>>=
 survdiff(Surv(duration, delta) ~ poverty, data = bfeed.short)
+@
+\normalsize
+\normalsize
+\label{fig:bfeed-poverty-test-short}
+\caption{Testing $H_0: S_{p = 1} = S_{p = 0}$ for data truncated at 36 weeks ($d \leq 36$)}
+\end{figure}
+
+\begin{figure}[h]
+\scriptsize
+<<bfeed-poverty-test-tail, echo = FALSE>>=
 survdiff(Surv(duration, delta) ~ poverty, data = bfeed.tail)
@@
 \normalsize
+\normalsize
+\label{fig:bfeed-poverty-test-tail}
+\caption{Testing $H_0: S_{p = 1} = S_{p = 0}$ for data truncated at 20 weeks ($d \geq 20$)}
+\end{figure}
 
